Interfacing logic programming and Coq

### 2020-10-22 papers/types2020/main.tex
Some progress on introduction and motivation.   We need to discuss.

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\input preamble
\input head

@@

Recently, C. Sacerdoti Coen and E. Tassi developed the ELPI
implementation \cite{dunchev15lpar} of \lP 
\cite{miller12proghol}.  More recently, Tassi has made ELPI available
as a plugin to the Coq proof assistant \cite{tassi18coqpl}.
%
This implementation of \lP extends earlier implementations in
primarily two directions.  First, ELPI adds a notion of contraints,
which makes it more expressive than, say, the Teyjus implementation of
\lP \cite{nadathur99cade}.  Second, the plugin version of ELPI comes
equip with an API for accessing the Coq environment and a quotation
and anti-quotation syntax allowing for Coq expressions to be mixed in
with \lP program elements.

ELPI provides a means for treating bindings and unification (of
structures containing bindings) that are convenient for various
meta-programming tasks in logic programming all things useful for
tactic programming in Coq (proof refinement).  For examples,
ELPI programmers are spared from dealing with low-level aspects of
representing bindings (e.g., De Bruijn indexes) while still having
clean and effective ways to manipulate binding structures.
%
In short, ELPI appears to be a useful meta-language for Coq
\cite{coen19mscs,tassi18coqpl,tassi19itp}.
%
In this paper, we outline two applications of the Coq plugin for ELPI
that support this claim.
%
With these examples, we shall illustrate that ELPI is a useful
meta-language not just because for its treatment of bindings and
syntactic structures in general but also because it is actually based
on a sound implementation of a higher-order intuitionistic logic.
%
Type inference is a well-known example of a meta-programming task for
which logic programming often provides immediate and elegant
implementations via the sound implementation of logic.
%
In this paper, we present two additional project in which an
implementation of proof search provides direct, elegant, and compact
implementations of meta-programming tasks.
%
Before we present these two examples, we first discuss the different
world views that the integration of ELPI and Coq focuses us to
confront.

@@

{\color{red} The rest of this section are just notes.}

These applications are not necessarily focused on the underlying
declarative nature of lambda Prolog.  For example, ELPI has various
extensions, such as ``spilling'' that are designed to make some
aspects of logic programming appear to be more functional.

In this paper, we limit our attention mostly to using lambda Prolog
for declarative processing: i.e., where logic plays a central role in
describing operations.  (Type checking is an example of such
processing, but we do not focus on type checking here).

Of course, being able to do declarative and not-so-declarative
operations within the same language makes that language all the more
useful.

  Given that lP and Coq are both examples of mixing the
  lambda-calculus with logic, it is important to understand their
  difference, something that can be very confusing.  Take the
  confusion around the term HOAS.  In some circles, this is a
  appropriate and sensible approach.  In other circles, it is
  semantically questionable.  But, in fact, this concept (using
  meta-level binding for object-level binding) plays our 
  differently depending on the meta language!

  The origin story to ML (Meta language) is that it supported
  LFC-style tactics and tacticals.  The strong typing of ML provided
  some guarantees: only theorems were ever given type thm.  But many
  steps of evolution have now taken place since then.  Isabelle's
  different approach to tactics.  Logic and proof theory, when used
  with Church's simple theory of types and Gentzen's sequent calculus,
  provides a far richer meta language: that provides an immediate and
  declarative treatment of bindings as well as for proof search
  (without the need of exception handling, etc).  [I'm making the case
  that FP is not a particularly natural framework for proof
  assistants.  Of course, lots of libraries have now been built in the
  FP style.]

  - Various dichotomies we find in this work.
    - proof theory vs type theory
    - logic programming vs functional programming
    - sequent calculus vs natural deduction

  - While both lambda Prolog and Coq are based on an amalgamation of
    $\lambda$-calculus and logic, these two languages take strikingly
    different approaches to bindings.  In lambda Prolog, bindings are
    syntactic and extensionality at any type is not available.  In
    Coq, however, higher-order typed quantification encodes
    extensional functions.  In lambda Prolog, (exists w. [lambda x. x]
    = [lambda x. w]) fails, no matter the type of w.  If this query
    was about functions, then this equation would be true when the
    type of w is a singleton.  But in lP, this is a query only about
    binding structures, and this fails to be proved since there is no
    capture-avoiding substitution (the only substitution supported by
    Church's STT) for w that makes the resulting two terms alpha,
    beta, eta convertible.  [Is this true in Coq?  What is equality
    here?]

  - On the proof theory side of this work, we use two different proof
    systems.

    - LJF, here we have undefined predicates, no induction
    - muMALL, were we have inductive definitions of predicates

    The Coq world provides for both of these kinds of predicate using
    Prop (corresponds to undefined) and Set (inductively defined).

@@

\section{Property based testing for Coq}

@@

% Not part of git repo.  Need to find a fix
\bibliography{../../../../repo/parsifal/papers/references/master}

\end{document}

>>>>

%% Important dates:
%%  Paper submission: 19 October 31 October 2020
%%  Author notification: 18 January 2021
%%  Final version:  15 February 2021
%%  Publication (presumably): end of March 2021

%% Style requirements of LIPIcs.  The recommended length of a paper is
%% 12-15 pages, excluding front-page(s) (authors, affiliation, keywords,
%% abstract, ...), bibliography and an appendix of max 5 pages. If you
%% need more pages, please ask the editors.

%% Papers have to be submitted in pdf through EasyChair:
%% https://easychair.org/conferences/?conf=types2020postproceed

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\input head
\input preamble

@@

Recently, C. Sacerdoti Coen and E. Tassi developed the ELPI
implementation \cite{dunchev15lpar} of \lP \cite{miller12proghol}.
More recently, Tassi has made ELPI available as a plugin to the Coq
proof assistant \cite{tassi18coqpl}.  This implementation of \lP
extends earlier implementations in primarily two directions.  First,
ELPI adds a notion of constraints, which makes it more expressive than,
say, the Teyjus implementation of \lP \cite{nadathur99cade}.  Second,
the plugin version of ELPI comes equip with an API for accessing the
Coq environment and a quotation and anti-quotation syntax allowing for
Coq expressions to be mixed in with \lP program elements.

ELPI provides a means for treating bindings and unification (of
structures containing bindings) that are convenient for various
meta-programming tasks in logic programming all things useful for
tactic programming in Coq (proof refinement).  For examples, ELPI
programmers are spared from dealing with low-level aspects of
representing bindings (e.g., De Bruijn indexes) while still having
clean and effective ways to manipulate binding structures.  In short,
ELPI appears to be a useful meta-language for Coq
\cite{coen19mscs,tassi18coqpl,tassi19itp}.  In this paper, we outline
two applications of the Coq plugin for ELPI that support this claim.
With these examples, we shall illustrate that ELPI is a useful
meta-language not just because for its treatment of bindings and
syntactic structures in general but also because it is actually based
on a sound implementation of a higher-order intuitionistic logic.
Type inference is a well-known example of a meta-programming task for
which logic programming often provides immediate and elegant
implementations via the sound implementation of logic.  In this paper,
we present two additional project in which an implementation of proof
search provides direct, elegant, and compact implementations of
meta-programming tasks.  Before we present these two examples, we
first discuss the different world views that the integration of ELPI
and Coq forces us to confront.

@@

In proof theory, one learns quickly that things always come in pairs:
negative/positive, left/right, bottom-up/top-down,
premises/conclusion, introduction/elimination,
% determinism/nondeterminism,
% invertible/noninvertible, 
etc.
%
When we examine the larger setting of this project of linking a logic
programming engine with Coq and its kernel, we find a large number of
pairing that are valuable to explicitly describe.

\subsection{Proof theory vs type theory}

In many ways, (structural) proof theory is more elementary and
low-level than most approaches to type theory.  For example, type
theories usually answer the question ``What is a proof?'' with the
response ``a dependently typed $\lambda$-term''.  That is, when
describing a type theory, one usually decides that a proof is a
certain kind of term within the system.  In contrast, the proof theory
treats logical propositions and proofs as separate.  For example,
proof theory does not assume that there are terms within the logic
that describe proofs.

Gentzen's discovery that the key to treating classical and
intuitionistic logics in the same proof system was the invention of
the weakening and contraction structural rules on the right-side of
sequents~\cite{gentzen35}.  This discovery lead him away from natural
deduction to the invention of multiple-conclusion sequent calculus, a
very important kind of proof that has not been given a satisfactory
representation in type theory.  This same innovation of Gentzen also
opened the way to the invention of another key proof theoretic
discovery, that of linear logic \cite{girard87tcs}.
Similarly, sequent calculus provides an elegant presentation for
linear logic while most treatments of linear logic in type theory
usually only capture linearity in a partial fashion.

Proof theory also has rather immediate and natural treatments for
co-induction (as well as induction: see
\cite{baelde12tocl,momigliano12jal}) whereas the treatment of
coinduction within type theory remains more challenging and is well
developed only for various syntactic conditions (see, for example,
\cite{bertot08entcs}).

\subsection{Proof search vs proof normalization}

Gentzen-style proofs are used to model computation in at least two
different ways.  The functional programming paradigm, following the
Curry-Howard correspondence, models computation abstractly as the
$\beta$-reduction of natural deduction proof: that is, computation is
modeled using \emph{proof normalization}.  On the other hand, the
logic programming paradigm, following the notion of goal-directed
proof search, models computation abstractly as a regimented search of
cut-free sequent calculus proofs: that is, computation is modeled
using \emph{proof search}.

Proof search has proved to have some richness that is hard to capture
in the proof normalization setting.  In particular, Gentzen's
invention of \emph{eigenvariables} are a kind of a proof-level
binder.  In the proof normalization setting, such eigenvariables
are instantiated during proof normalization.  In contrast, during the
search for cut-free proofs, eigenvariables are part of the syntax of
terms and formulas.  As a result, they can be used in the treatment of
bindings in data structures more generally.  Hence, the logic
programming paradigm provides a completely natural, elegant, and
powerful method, called $\lambda$-tree syntax, for treating bindings
within syntax.

In this paper, we shall use two different proof systems.  The first
logic is first-order intuitionistic logic.  Here, predicates
are not inductively defined and quantification will be limited to
range over first-order term.  Actually, we shall examine only the Horn
clause fragment of that logic.
%
The second logic will allow for inductively defined predicates using a
fixed point notation.   Given again our restriction to essentially
Horn clause specifications, it is possible to view intuitionistic
reasoning with such specification as if they are actually
specification in $\mu$MALL, an extension of multiplicative additive
linear logic with the least and greatest fixed point operators
\cite{baelde12tocl,baelde07lpar,heath19jar}.
%
The difference between these two logics and associate proof theories
is reminiscent to the difference between the type \lsti{Prop}
(corresponding to undefined propositions) and Sets (corresponding to
inductively defined propositions).

\subsection{\lP vs Coq}

Given that \lP and Coq are both examples of mixing the
$\lambda$-calculus with logic, it is important to understand their
differences.  The confusion around the term \emph{higher-order
abstract syntax} (HOAS), for example, helps to illustrate their
differences.  In the functional programming setting, including the Coq
system, the HOAS approach leads to encoding binding structures within
terms using functions.  Such encodings have results in surprisingly
complicated encodings, allowing, for example, for \emph{exotic terms}
\cite{despeyroux95tlca} and to structures on which induction is not
possible \cite{roeckl01fossacs}.  In the logic programming setting,
particularly \lP, HOAS is well supported in both the availability of
binders allowed with terms ($\lambda$-bindings), in formulas
(quantifiers), and in proofs (eigenvariables).  For this reason, the
term $\lambda$-tree syntax was introduced to name this particular take
on HOAS \cite{miller19jar}.  The Abella proof assistant
\cite{baelde14jfr} was designed, in part, to allow for inductive and
co-inductive proofs for specifications using the $\lambda$-tree syntax
approach.

Another difference between \lP and functional programming can be
illustrated by considering how they are used in the specification of
tactics that are popular in modern proof assistants.  In fact, the
origin story for the ML functional programming language was that ML
was the meta-language for implementing the LCF suite of tactics and
tacticals \cite{gordon79}.  In order to implement tactics, ML adopted
not only novel features such as polymorphically typed higher-order
functional programming but also the non-functional mechanisms of
failure and exception handling.  While \lP is also based on
polymorphically typed higher-order relational programming, it also
comes with a completely declarative version of failure and
backtracking search.  Combining those features along with its support
of unification (even in the presence of term-level bindings), \lP
provides a rather different approach to the specification of tactics
\cite{felty93jar}.

While \lP is a typed language, it only uses simple types and these are
more often used to indicate syntactic categories.  Any dependency
information needs to be captured using predicates and quantifiers.

\section{Proof theory and proof certificates}

{\color{red} The goal in this section is to say a bit about focusing
  generally, and to give a bit of detail about FPC for Horn clauses.
  Initial draft of text taken from \cite{blanco19ppdp}.
}

\subsection{Focused proof systems}
\label{ssec:focused}

The proof search approach to encoding Horn clause computation
results in the structuring of proofs with repeated
switching between a \emph{goal-reduction} phase and a
\emph{backchaining} phase~\cite{miller91apal}.
%
The notion of \emph{focused proof systems}~\cite{andreoli92jlc,liang09tcs}
generalizes this view of proof construction by identifying the
following two phases. % of proof construction.
\begin{enumerate}
\item The \emph{negative}%
\footnote{The terminology of negative and positive phases is a bit
  unfortunate: historically, these terms do not refer to positive or negative
  subformula occurrences but rather to certain semantic models used in
  the study of linear logic~\cite{girard87tcs}.}
%
  phase corresponds to  goal-reduction: in this phase,
  inference rules that involve \emph{don't-care-non\-de\-ter\-min\-ism} are
  applied.
%
  As a result, there is no need to consider backtracking over choices
  made in building this phase.

\item The \emph{positive} phase corresponding to  backchaining: in this phase, inference rules that involve
  \emph{don't-know-non\-de\-ter\-min\-ism} are applied: here,
  inference rules need to be supplied with information in order to
  ensure that a completed proof can be built.
%
  That information can be items such as which term is needed to
  instantiate a universally quantified formula and which disjunct of a
  disjunctive goal formula should be proved.
\end{enumerate}
%
Thus, when building a proof tree (in the sequent calculus) from the
conclusion to its leaves, the negative phase corresponds to a simple
computation that needs no external information, while the positive
phase may need such external information to be supplied.
%
In the literature, it is common to refer to a repository of such
external information as either an \emph{oracle} or a \emph{proof
certificate}.

When using a focused proof system for logic extended with fixed
points, such as  employed in Bedwyr~\cite{baelde07cade} and described
in~\cite{baelde12tocl,heath17linearity}, proofs of formulas such as
\[
  \exists x [(\tau(x)\wedge P(x)) \wedge \neg Q(x)]
  \tag{*}\label{eq:full}
\]
are a single \emph{bipole}: that is, when reading a proof bottom up, a
positive phase is followed on all its premises by a single negative
phase that completes the proof.%
\footnote{The reader familiar with focusing will understand that there
are two ``polarized'' conjunctions, written in linear logic as
$\otimes$ and $\with$ or in classical and intuitionistic logics as
$\wedgep$ and $\wedgen$, respectively.  In this paper, we use simply
$\wedge$ to denote the positive biased conjunction.}
%
In particular, the positive phase corresponds to the \emph{generation} phase
and the negative phase corresponds to the \emph{testing} phase.
%

Instead of giving a full focused proof system of a logic including
fixed points (since, as we will argue, that proof system will not, in
fact, be needed to account for PBT ), we offer the following analogy.
%
Suppose that we are given a finite search tree and we are asked to
prove that there is a secret located in one of the nodes of that
tree.
%
A proof that we have found that secret can be taken to be a
description of the path to that node from the root of the tree: that
path can be seen as the proof certificate for that claim.
%
On the other hand, a proof that no node contains the secret is a
rather different thing: here, one expects to use a blind and
exhaustive search (via, say, depth-first or breath-first search) and
that the result of that search never discovers the secret.
%
A proof of this fact requires \emph{no} external information: instead it
requires a lot of computation involved with exploring the tree until
exhaustion.
%
This follows the familiar pattern where the positive (generate) phase
requires external information while the negative (testing) phase
requires none.
%
Thus, a proof certificate for $(\ref{eq:full})$ is also a proof
certificate for 
\[
  \exists x [\tau(x)\wedge P(x)].
  \tag{**}\label{eq:short}
\]
Such a certificate would contain the witness (closed) term $t$ for the
existential quantifier and sufficient information to confirm that
$P(t)$ can be proved (a proof of a typing judgment such as $\tau(t)$
is usually trivial).
%
%In short\ednote{AM: a bit repetitious}, a proof certificate for
%$(\ref{eq:short})$ can play the part 
%of a proof certificate also for $(\ref{eq:full})$, if we also use
%negation-as-finite-failure  with $Q(t)$.
%  ****
Since a proof certificate for the existence-of-a-counterexample formula
$(\ref{eq:full})$ can be taken as a proof certificate of
$(\ref{eq:short})$, then we only need to consider proof certificates
for Horn clause programs.
%
%Since the logic of Horn clauses is rather simple,
We illustrate next what such proofs and proof certificates look like
for the rather simple logic of Horn clauses.

\subsection{Certificate checking with expert predicates}
%\subsection{Adding certificates and experts to a proof system}
\label{ssec:fpc}

%% \begin{metanote}
%%   AM: I know this is well known and boring, but using \texttt{prog A
%%     G} in Fig \ref{fig:augmented} is imprecise --- in the comment of
%%   next figure we mention \texttt{A :- G} and we gloss over the
%%   instantiation of \texttt{prog A' G} to match the goal \texttt{A},
%%   which we assume to be ground, given the exists rule
%% \end{metanote}

%% DM I've modified the text a bit to hopefully improve the state of
%% things here.  The formal proof theory is all about closed
%% formulas.  The prolog implementation allows open formulas.  This
%% difference invades most aspects of proof checking, not just the
%% prog predicate.

% DM We should find a place to smooth over this issue (which is not
% hard, I think).

\begin{figure}
\newcommand{\XXi}{{\color{blue}{\Xi}}}
\[
\infer{\XXi\vdash t = t}
      {\eqExpert{\XXi}}
\qquad
\infer{\XXi\vdash \true}
      {\trueExpert{\XXi}}
\qquad
\infer{\XXi\vdash G_1\vee G_2}
      {\XXi'\vdash G_i\qquad \orExpert{\XXi}{\XXi'}{i}}
\qquad
\infer{\XXi\vdash \exists x. G}
      {\XXi'\vdash G[t/x]\qquad \someExpert{\XXi}{\XXi'}{t}}
\]
\vskip -6pt
\[
\infer{\XXi\vdash G_1\wedge G_2}
      {\XXi_1\vdash G_1\quad \XXi_2\vdash G_2\quad \andExpert{\XXi}{\XXi_1}{\XXi_2}}
\qquad
\infer{\XXi\vdash A}
      {\XXi'\vdash G \quad (A~\hbox{\tt :-}~G)\in\instan\Pscr
                     \quad \unfoldExpert{\XXi}{\XXi'}}
\]
\caption{A proof system augmented with proof certificates and
  additional ``expert'' premises.}
\label{fig:augmented}
\lstinputlisting[linerange={sigs-end}]{code/kernel.sig}
\lstinputlisting[linerange={check-end}]{code/kernel.mod}
\caption{A simple proof checking kernel.}
\label{fig:kernel}
\end{figure}

Figure~\ref{fig:augmented} contains a simple proof system for Horn
clause provability in which each inference rule is augmented with an
additional premise involving an \emph{expert predicate}, a
certificate $\Xi$, and possibly continuations of certificates ($\Xi'$,
$\Xi_1$, $\Xi_2$) with extracted information from certificates (in the
case of $\vee$ and $\exists$).
%
%
The premise $(A~\hbox{\tt :-}~G)\in\instan{\Pscr}$ in the last
inference rule of Figure~\ref{fig:augmented} states that the logic
programming clause $(A~\hbox{\tt :-}~G)$ is a ground instance of some
clause in a fixed program $\Pscr$.
%
Although this proof system is a focused proof system, the
richness of focusing is not apparent in this simplified 
setting: thus, we drop the adjective ``focused'' from this point forward.

Figure~\ref{fig:kernel} contains the \lP implementation of the
inference rules in Figure~\ref{fig:augmented}: here the infix
turnstile $\vdash$ symbol is replaced by the \lst<i{check} predicate.
%
Notice that it is easy to show that no matter how the expert predicates
are defined, if the goal \lsti{check Cert B} is provable in \lP then
\lsti{B} must be a sound consequence of the program clauses stored in
the \lsti{prog} predicate (which provides a natural implementation of
the premise $(A~\hbox{\tt :-}~G)\in\instan{\Pscr}$.
% AM: not there
% in Figure~\ref{fig:kernel}).

As we mentioned in the introduction, our notion of proof certificates
 is taken from the general setting of \emph{foundational
  proof certificates}~\cite{chihani17jar}.
%
In our case here, an FPC is a collection of \lP clauses that
provide the remaining details not supplied in Figure~\ref{fig:kernel}:
that is, the exact set of constructors for the \lsti{cert} type as
well as the exact specification of the six expert predicates listed in
that figure.
%
Figure~\ref{fig:resources} displays two such FPCs,
both of which can be used to describe proofs where we bound
the number of occurrences of unfoldings in a proof.
%
For example, the first FPC
provides the experts for treating certificates that are constructed
using the \lsti{qheight} constructor.
%
As is easy to verify, the query \mbox{\lsti{(check (qheight 5) B)}} (for the
encoding \lsti{B} of a goal formula) is provable in \lP using the
clauses in Figures~\ref{fig:kernel} and~\ref{fig:resources} if and
only if the height of that proof is 5 or less.
%
Similarly, the second FPC uses the constructor \lsti{qsize} (with two
integers) and can be used to bound the total number of instances of
unfoldings in a proof.
%
In particular, the query \mbox{\lsti{sigma H\ (check (qsize 5 H) B)}}
% \begin{lstlisting}
% sigma H\ (check (qsize 5 H) B)
% \end{lstlisting}
is provable if and only if the total number of unfoldings of that
proof is 5 or less.

\begin{figure}
\lstinputlisting[linerange={resources-end}]{code/fpcs.sig}
\lstinputlisting[linerange={resources-end}]{code/fpcs.mod}
\caption{Two FPCs that describe proofs that are limited in either
  height or in size.}
\label{fig:resources}
\end{figure}

Finally, Figure~\ref{fig:max} contains the FPC based on the constructor
\lsti{max} that is used to record explicitly all information within a
proof: in particular, all disjunctive choices and all substitution
instances for existential quantifiers are collected into a binary tree
structure of type \lsti{max}.
%
In this sense, proof certificates built with the \lsti{max}
constructor are \emph{maximally} explicit.
%
Such proof certificates are used, for example, in~\cite{Pair}; it is
important to note that proof checking with such maximally explicit
certificates can be done with much simpler proof-checkers than those
used in logic programming since backtracking search and unification
are not needed.
%

\begin{figure}
\lstinputlisting[linerange={max-end}]{code/fpcs.sig}
\lstinputlisting[linerange={max-end}]{code/fpcs.mod}
\caption{The \lsti{max} FPC}
\label{fig:max}
\end{figure}

Further, if we view a particular FPC as a means of restricting proofs, it is
possible to build an FPC that \emph{restricts} proofs satisfying two FPCs
simultaneously.
%
In particular, Figure~\ref{fig:pairing} defines an FPC based on the
(infix) constructor \lsti{<c>}, which \emph{pairs} two terms of type
\lsti{cert}.
%
The pairing experts for the certificate \lsti{Cert1 <c> Cert2} simply
request that the corresponding experts also succeed for both
\lsti{Cert1} and \lsti{Cert2} and, in the case of the \lsti{orE} and
\lsti{someE}, also return the same choice and substitution term,
respectively.
%
Thus, the query 
\begin{lstlisting}
?- check ((qheight 4) <c> (qsize 10)) B
\end{lstlisting}
will succeed if there is a proof of \lsti{B} that has a height less
than or equal to 4 while also being of size less than or equal to 10.
%
A related use of the pairing of two proof certificates is to %use it to
\emph{distill} or \emph{elaborate} proof certificates. 
%
For example, the proof certificate \lsti{(qsize 5 0)} is rather implicit
since it will match any proof that used unfold exactly 5 times.
%
However, the query
\begin{lstlisting}
?- check ((qsize 5 0) <c> (max Max)) B.
\end{lstlisting}
will store into the \lP variable \lsti{Max} more complete details of
any proof that satisfies the \lsti{(qsize 5 0)} constraint.
%
In particular, this forms the infrastructure of 
 an \emph{explanation} tool for attributing ``blame'' for
the origin of a counterexample; these maximal
certificates are an appropriate starting point for documenting
both the counterexample and why it serves as a counterexample.

\begin{figure}
\lstinputlisting[linerange={pairing-end}]{code/fpcs.sig}
\lstinputlisting[linerange={pairing-end}]{code/fpcs.mod}
\caption{FPC for pairing}
\label{fig:pairing}
\end{figure}

Various  additional examples and experiments using the pairing of FPCs
can be found in~\cite{Pair}. Using similar techniques, it is possible to define FPCs that target
specific types for special treatment: for example, when generating
integers, only (user-defined) small integers could be inserted into
counterexamples.

\section{Property-based testing for Coq}

Simplest setting since we are only sending back to Coq a term (a
counterexample), not a general proof.

@@

@@

\bibliography{l}
\end{document}
